Verifying synchronous distributed algorithms using IVy

### 2018-08-12 paper/paper.tex
Minor fixes

\section{IVy}
IVy is a system designed to verify correctness of infinite-sate systems using the aforementioned deductive approach. It takes as input a specification of an algorithm operating on system-state  and a set of invariants. It proves the inductiveness of the invariants in respect to the algorithm, or displays a counterexample to induction. If the set of invariants entails the desired properties of the algorithm, and Ivy proved induction, then it was proven that the algorithm is \textit{correct}.

@@

A detailed discussion about decidability lies outside of the scope of this work and is well documented in \cite{decide}.

@@

The assumption $f<n$, ``number of failed nodes is bounded to an amount smaller than the number of total nodes'', can alternatively be expressed via $\exists n\ \neg failed(N)$. This is because the existence of a functioning node implies that the number of failed nodes $f$ is smaller than $n$.

@@

\subsection{Step 1 - Message sending}
The contents of a node's message is modeled via the relation $message(N:node,V:value)$. Where
$message(1,0) = true; message(1,1) = true;$ means node $1$ includes in its message the values $0$ and $1$.
Since a node sends the same message to all nodes, for any given round, we do not need to include the
recipient in the relation.

The actual delivery of a message is modeled via the $incoming\_message(RECIPIENT:node,SENDER:node)$ relation. Where $incoming\_message(0,1) = true; incoming\_message(0,2) = true;$ means that node $0$ receives the messages from node $1$ and node $2$.

@@

Line 2 in the \textit{step2} action adds to $w$ of node $N$ all values $V$ for which a node $J$ exists that has sent a message to $N$ that contains $J$.
The conditional statement in Line 4 corresponds to the \textit{decision condition} ($if\ rounds = f+1$) in the original algorithm. Nodes with a single value in $w$ decide on that value, otherwise
they decide on the default value.
Finally, for good measure, $incoming\_message$ and $message$ are reset, but this is not necessary for the proof.

@@

Line 1 prepares for each node $M$ a non-deterministically chosen set of nodes that it sends a message to. This set is represented by the relation \textit{incoming\_message\_failing}. This relation can be thought of as the set of nodes $M$ will send a message to if $M$ fails.
Line 2 updates the \textit{incoming\_message} relation. If node $M$ failed during this round, the set of nodes
 $M$ sends messages to is changed to the values of \textit{incoming\_message\_failing}, otherwise it remains the same. It is important to point out that this approach correctly simulates the fact that nodes are not thought of as sending messages sequentially but in non-deterministic order. The specification makes no assumptions about an order of message sending. Furthermore, for fixed $m$, \textit{incoming\_message\_failing(N,m)} may still be $true$ for all N. The specification therefore includes the possibility of a node failing after it sent all its messages successfully.

@@

Now that the system has been modeled, the next step of verifying correctness with IVy is specifying inductive invariants. The simplest invariants to come up with are those, that describe the correctness of the algorithm at hand. In our case, the correctness is described by the \textit{correctness conditions} in \textit{iii}. These are \textit{agreement}, \textit{validity} and \textit{termination}.

@@

# if a value is in a process' W, it must be initial value of
# some other process (no invented values)

invariant node.w(N, V) -> (exists M . node.val(M)=V)

# lemma 1:
invariant system.clean_round_occured ->
            (~node.failed(A) & ~node.failed(B)) ->
              (node.w(A, V) = node.w(B, V)))
\end{lstlisting}
\end{mdframed}

@@

As these two lines are so different, equivalence is certainly not obvious to a user. Proving equivalence of pseudo-code and Ivy language specification would certainly be a non-trivial proof exercise that certainly trumps manually proving \textit{Floodset}'s correctness in difficulty.

\subsubsection{Lack of modularity}
Furthermore, it has already been shown in section \textbf{IV. i.1} that the lack of easy iteration over types makes it difficult to achieve modularity. The lack of which introduces additional uncertainty about correct modelling, especially in distributed systems.

@@

\subsection{Necessity of deep understanding}
It has become apparent, that successful verification with IVy is not possible without deep knowledge about how the algorithm functions. Ivy is not a ``translate and go'' solution.
Evidence of this include the complex invariant that requires knowledge about \textit{Lemma 1} or the alternative encoding for $f<n$ and the decision condition $if\ rounds = f+1$ that had to be discovered.

@@

\section{Conclusion}
This paper has presented the Ivy system and important Ivy language features. It classified IVy
as part of a three step process of algorithm verification.
It explained that its usage requires a readiness to let go of procedural programming paradigms
in favor of the logic oriented language. It gave an explanation of induction and inductive invariants,
explaining what they are, how they are used by IVy to prove correctness and how they can be found.
It presented some of the systems capabilities to verify distributed algorithms by example.

It pointed out some limitations. One being that it is hard to know an algorithm was correctly specified in IVy. This is due to
the lack of easy type iteration, which leads to complicated global assignments and a non-modular structure that is far off
the original algorithm's.
The paper suggested a mechanism to verify the possibility of a state to occur, to alleviate this uncertainty.
Another limitation that was found, is the difference between a logically sound algorithm and a correctly implemented one, and how IVy can only verify the former.
It was stressed that successful verification of an algorithm that is not thoroughly understood by the user is unlikely.

>>>>

\section{IVy}
IVy is a system designed to verify correctness of infinite-sate systems using the aforementioned deductive approach. In a nutshell, it takes as input the specification of an algorithm, which may mutate state, and a set of invariants. It proves the inductiveness of the invariants in respect to the algorithm, or displays a counterexample to induction. If the set of invariants entails the desired properties of the algorithm, and Ivy proved induction, then it was proven that the algorithm is \textit{correct}.

@@

A detailed discussion about decidability lies outside of the scope of this work and is well documented in \cite{decid}.

@@

The assumption $f<n$, ``number of failed nodes is bounded to an amount smaller than the total number of nodes'', can alternatively be expressed via $\exists n\ \neg failed(n)$. This is because the existence of a functioning node implies that the number of failed nodes $f$ is smaller than $n$.

@@

\subsection{Step 1 - Message sending}
The contents of a node's message is modeled via the relation $message(N:node,V:value)$. Where
$message(1,0) = true;\ message(1,1) = true;$ means node $1$ includes in its message the values $0$ and $1$.
Since a node sends the same message to all nodes, for any given round, we do not need to include the
recipient in the relation.

The actual delivery of a message is modeled via the \\
$incoming\_message(RECIPIENT:node,SENDER:node)$ relation. Where $incoming\_message(0,1) = true;\ incoming\_message(0,2) = true;$ means that node $0$ receives the messages from node $1$ and node $2$.

@@

Line 2 in the \textit{step2} action adds to $w$ of node $N$ all values $V$ for which a node $J$ exists that has sent a message to $N$ that contains $V$.
The conditional statement in Line 7 corresponds to the \textit{decision condition} ($if\ rounds = f+1$) in the original algorithm. Line 9 sets the helper relation $has\_single\_value(N:node)$ that is set to true for $N$ iff it has exactly one value in its $w$. Line 13 specifies that nodes with a single value in $w$ decide on that value, unless they have failed. Line 16 specifies that nodes with multiple values in their $w$ decide on their default value, unless they failed. The disjunction on line 17 with $node . decision (N, node . default (N))$ is necessary so as not to overwrite an already made decision of a node with $w = \{v_0\}$, $v_0$ being the default value.
Finally, for good measure, $incoming\_message$ and $message$ are reset, but this is not necessary for the proof.

@@

Line 1 prepares for each node $M$ a non-deterministically chosen set of nodes that it sends a message to. This set is represented by the relation \textit{incoming\_message\_failing}. This relation can be thought of as the set of nodes which $M$ will send a message to if $M$ fails.
Line 2 updates the \textit{incoming\_message} relation. If node $M$ failed during this round, the set of nodes which
 $M$ sends messages to is changed to the values of \textit{incoming\_message\_failing}, otherwise it remains the same. It is important to point out that this approach correctly simulates the fact that nodes are not thought of as sending messages sequentially but in non-deterministic order. The specification makes no assumptions about an order of message sending. Furthermore, for fixed $m$, \textit{incoming\_message\_failing(N,m)} may still be $true$ for all N. The specification therefore includes the possibility of a node failing after it sent all its messages successfully.

@@

Now that the system has been modeled, the next step of verifying correctness with IVy is specifying inductive invariants. The simplest invariants to come up with are those, that describe the correctness of the algorithm at hand. In our case, the correctness is described by the \textit{correctness conditions} in \textit{III.iii}. These are \textit{agreement}, \textit{validity} and \textit{termination}.

@@

# if a value is in a process' W, it must be initial value of
# some other process (no invented values)
invariant node.w(N, V) -> (exists M . node.val(M)=V)

# lemma 1, explained in III
invariant system.clean_round_occured ->
            (~node.failed(A) & ~node.failed(B)) ->
              (node.w(A, V) = node.w(B, V)))
\end{lstlisting}
\end{mdframed}

@@

As these two lines are so different, equivalence may not be obvious to a user. Proving equivalence of pseudo-code and Ivy language specification would be a non-trivial proof exercise that certainly trumps manually proving \textit{Floodset}'s correctness in difficulty.

\subsubsection{Lack of modularity}
Furthermore, it has already been shown in section \textit{IV. i.1} that the lack of easy iteration over types makes it difficult to achieve modularity. The lack of which introduces additional uncertainty about correct modelling, especially in distributed systems.

@@

\subsection{Necessity of deep understanding}
It has become apparent, that successful verification with IVy is not possible without deep knowledge about how the algorithm functions. Ivy is not a ``translate and go'' solution.
Evidence of this include the complex invariant that requires knowledge about \textit{lemma 1} or the alternative encoding for $f<n$ and the decision condition $if\ rounds = f+1$.

@@

\section{Conclusion}
This paper has presented the Ivy system and important Ivy language features. It classified IVy
as part of a two step process of algorithm verification.
It explained that its usage requires a readiness to let go of procedural programming paradigms
in favor of the logic oriented language. It gave an explanation of induction and inductive invariants,
explaining what they are, how they are used by IVy to prove correctness and how they can be found.
It presented some of the systems capabilities to verify distributed algorithms by example.

It pointed out some limitations. One being that it is hard to know an algorithm was correctly translated into the IVy language, as a one-to-one translation from a procedural algorithm is unlikely. This is due to the lack of easy type iteration, which leads to complicated global assignments and a non-modular structure that is far off the original algorithm's.
Furthermore, the system context may easily be miss-specified. The paper suggested a mechanism to verify the possibility of a state to occur, to alleviate this problem.
Another limitation that was found, is the difference between a logically sound algorithm and a correctly implemented one, and how IVy can only verify the former.
It was stressed that successful verification of an algorithm that is not thoroughly understood by the user is unlikely.
